Disparate views of disparity

### 2017-12-08 Analysis/disparity_methods.Rmd
PCO plots

```{r}
source("../Functions/fuzzy.sort.data.R")
```

@@

```{r}
## Reading the data
data_raw <- read.csv("../Data/disparity_methods.csv", stringsAsFactors = FALSE)
search_results <- read.csv("../Data/GoogleSearch_occurences.csv", stringsAsFactors = FALSE)

## Keeping only the studies with data
data <- data_raw[apply(data_raw[, 3:9], 1, function(X) ifelse(all(X == "-"), FALSE, TRUE)), ]
```

@@

```{r, fig.width = 8, fig.height = 8}
disparity_fuzzy <- search_results[-c(1:2, 60:65),3]
disparity_exact <- search_results[-c(1:2, 60:65),2]

## Number of papers
# pdf("../Paper/Figures/GoogleScholarOccurences.pdf", width = 8, height = 8)
op <- par(bty = "n")
plot(seq(from = 1960, to = 2016), rev(disparity_fuzzy)/search_results[1,3], type = "l", xlab = "", ylab = "Relative number of papers per year", ylim = c(0, 0.1))
lines(seq(from = 1960, to = 2016), rev(disparity_exact)/search_results[1,2], type = "l", col = "grey")
par(op)
# dev.off()
```

@@

## Sorting the variables

The variables can be sorted as follow:

```{r}

## Categories definition
data_type <- c("3D landmarks", "2D landmarks", "continuous", "discrete", "body", "metabolic")
full_organism <- c("yes", "no")
morphospace <- c("PCA", "MDS", "PCO", "PCOA", "Raup")
disparity <- c("sum of variance", "sum of range", "product of variance", "product of range", "distance", "observed") #TG: Area from observed to estimated (DTT)
test <- c("anova", "model fitting", "canonical", "confidence interval")
phylo <- c("yes", "no", "mapped")

factors <- list("data type" = data_type, "full organism" = full_organism, "morphospace" = morphospace, "disparity" = disparity, "test" = test, "phylo correction" = phylo)
```

```{r, fig.width = 12, fig.height = 12}
## Plotting the results
par(mfrow = c(2, 3))
for(factor in 1:length(factors)) {
    plot.data(fuzzy.sort.data(data[,3+factor], factors[[factor]]), proportion = TRUE, ylim = c(0, 100), main = names(factors)[[factor]])    
}
```

```{r}
## Printing the tables
factor <- 1
table.data(fuzzy.sort.data(data[,3+factor], factors[[factor]]), table = "tex")

@@

@@

```{r}
disparity_fuzzy <- search_results[-c(1:2, 60:65),3]
disparity_exact <- search_results[-c(1:2, 60:65),2]

## Number of papers
pdf("../Writing/Figures/GoogleScholarOccurences.pdf", width = 8, height = 8)
op <- par(bty = "n")
plot(seq(from = 1960, to = 2016), rev(disparity_fuzzy)/search_results[1,3], type = "l", xlab = "", ylab = "Number of papers per year", ylim = c(0, 0.1), yaxt = "n")
lines(seq(from = 1960, to = 2016), rev(disparity_exact)/search_results[1,2], type = "l", col = "grey")
## Adding the axis
axis(2, at = seq(from = 0, to = 0.08, by = 0.02), labels = round(seq(from = min(disparity_fuzzy), to = max(disparity_fuzzy), length.out = 5)) )
axis(4, at = seq(from = 0, to = 0.1, by = 0.02), labels = round(seq(from = min(disparity_exact), to = max(disparity_exact), length.out = 6)), col = "grey", las = 2)

par(op)
dev.off()
```

@@

plot(add, type = "line", xlab = "Number of Google Scholar matches", ylab = "Number of new entries")

>>>>

```{r}
source("../Functions/fuzzy.sort.data.R")
source("../Functions/plot.data.R")
```

@@

```{r}
## Reading the data
cleaned_data <- readRDS("../Data/Processed/cleaned_data.Rda")
search_results <- read.csv("../Data/GoogleSearch_occurences.csv", stringsAsFactors = FALSE)

## Extracting the cleaned data
factors <- cleaned_data$factors
data <- cleaned_data$data
```

@@

```{r, fig.width = 8, fig.height = 8}
## Selecting the right data
disparity_fuzzy <- search_results[-c(1:2, 60:65),3]
disparity_exact <- search_results[-c(1:2, 60:65),2]

## Number of papers
# pdf("../Writing/Figures/GoogleScholarOccurences.pdf", width = 8, height = 8)
op <- par(bty = "n")
plot(seq(from = 1960, to = 2016), rev(disparity_fuzzy)/search_results[1,3], type = "l", xlab = "", ylab = "Number of papers per year", ylim = c(0, 0.1), yaxt = "n")
lines(seq(from = 1960, to = 2016), rev(disparity_exact)/search_results[1,2], type = "l", col = "grey")
## Adding the axis
axis(2, at = seq(from = 0, to = 0.08, by = 0.02), labels = round(seq(from = min(disparity_fuzzy), to = max(disparity_fuzzy), length.out = 5)) )
axis(4, at = seq(from = 0, to = 0.1, by = 0.02), labels = round(seq(from = min(disparity_exact), to = max(disparity_exact), length.out = 6)), col = "grey", las = 2)
par(op)
# dev.off()
```

@@

## Variable distribution

What is the proportion of use for each method?

```{r, fig.width = 18, fig.height = 6}
## Plotting the results
pdf("../Writing/Figures/MethodsProportions.pdf", width = 18, height = 6)
op <- par(mfrow = c(2, 3), bty = "n")
for(factor in 2:length(factors)) {
    plot.data(fuzzy.sort.data(data[,factor], factors[[factor]]), proportion = TRUE, ylim = c(0, 100), main = names(factors)[[factor]])
}
par(op)
dev.off()
```

Methods seem to full in three main categories:

  * The `Claddis` approach: Discrete data for the full organism ordinated with a PCO with a variance based disparity metric and no phylogenetic correction.
  * The `geomorph` approach: landmark data for part of the organism ordinated with a PCA with a distance based disparity metric and no phylogenetic correction.
  * The `dtt` approach: continuous data for the full organism, not ordinated, with a pairwise distance based disparity metric and phylogenetic correction.

```{r}
## Methods types table (empty)
methods_types <- matrix(NA, nrow = 3, ncol = 6, dimnames = list(c("Claddis", "geomorph", "dtt"), c(names(factors)[-1])))

## Filling the table
methods_types[,1] <- c("discrete", "landmarks", "continuous")
methods_types[,2] <- c("yes", "no", "yes")
methods_types[,3] <- c("PCO", "PCA", "\"pairwise matrix\"")
methods_types[,4] <- c("variance", "distance", "pairwise distance")
methods_types[,5] <- c("variable", "anova", "models")
methods_types[,6] <- c("no", "sometimes", "yes")

knitr::kable(methods_types)
```

## A disparity-space?

Multivariate data? Let's analysis it using disparity!

But first, a bit of cleaning

```{r}
## Removing the taxa column
multi_data <- apply(data[,-1], 2, as.character)

## Replace NAs by "-"
multi_data <- ifelse(is.na(multi_data), "-", multi_data)

## Selecting only the first variable when multiples
multi_data <- gsub(" \\+ .*", "", multi_data)

## Transform the data as numeric
multi_data <- as.list(as.data.frame(multi_data))
multi_data <- lapply(multi_data, as.numeric)
multi_data <- matrix(unlist(multi_data), ncol = ncol(data[,-1]), byrow = FALSE)

## Make a pairwise matrix
pairwise_data <- vegan::vegdist(multi_data, method = "bray")

## Ordinating the distance matrix
ordinated_data <- stats::cmdscale(pairwise_data, k = sum(unlist(lapply(factors, length))[-1]) - 2)
#TG: Axis are selected as the number of factors - 2
```

```{r, fig.width = 8, fig.height = 8}
## Get the variance per axis
scree_plot <- dispRity::variances(ordinated_data)
percentages_scree <- round(scree_plot/sum(scree_plot)*100, digit = 2)

## Variance per dimensions
barplot(scree_plot, ylab = "Variance per dimension", xlab = "Dimensions")
```

```{r, fig.width = 8, fig.height = 8}

factor <- 5
# length(unique(multi_data[,1]))
# length(factors[[2]])

col_palette <- rainbow(length(unique(data[, factor])))
colours <- col_palette[as.factor(multi_data[, factor-1])]

plot(ordinated_data[,1], ordinated_data[,2], xlab = paste0("PC1 (",percentages_scree[1], "%)"), ylab = paste0("PC2 (",percentages_scree[2], "%)"), pch = 19, col = colours, main = paste0(names(factors)[factor]))

legend("bottomright", legend = sort(factors[[factor]]), col = col_palette, pch = 19, cex = 0.5)

@@

<!-- ```{r}
## Printing the tables
factor <- 1
table.data(fuzzy.sort.data(data[,factor], factors[[factor]]), table = "tex")

``` -->

@@

## Testing the three categories hypothesis

Setting the data for the three categories hypothesis:

  * The `Claddis` approach: Discrete data for the full organism ordinated with a PCO with a variance based disparity metric and no phylogenetic correction.
  * The `geomorph` approach: landmark data for part of the organism ordinated with a PCA with a distance based disparity metric and no phylogenetic correction.
  * The `dtt` approach: continuous data for the full organism, not ordinated, with a pairwise distance based disparity metric and phylogenetic correction.

```{r}
cleaned_data <- readRDS("../Data/Processed/cleaned_data.Rda")

data <- cleaned_data$data

@@

plot(add, type = "line", xlab = "Number of Google Scholar matches", ylab = "Number of new entries")
```

## Groups studied

Which groups are studied using disparity analysis?

```{r}
plot.data(fuzzy.sort.data(data[,1], factors[[1]]), proportion = TRUE, ylim = c(0, 100), main = names(factors)[[1]])
```

It's pretty much even with no big majority!
